---
title: "5 Regression with more than one predictor"
author:
- first_name: Thomas
  last_name: Brochhagen
  url: https://tbrochhagen.github.io/
  affiliation: Universitat Pompeu Fabra
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
  pdf_document:
    toc: yes
bibliography: ../../biblio.bib
---



```{r preamble, echo=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(ggpubr)
```

## Regression with multiple predictors

We can straightforwardly extend our linear model with a single predictors to account for multiple ones:

$$y = \beta_0 + \beta_1 x_1 + ... + \beta_n x_n + \text{error} $$

In words, outcome $y$ is modelled as a linear combination of $n+1$ parameters, $\beta_0$,  $\beta_1$ up to $\beta_n$, and an error term. The contribution of $\beta_1$ to the prediction is relative to the value of $x_1$, that of $\beta_2$ is relative to the value of predictor $x_2$ and so on. 

#### Case study: pitch (continued)

You already built a regression model with multiple predictors in the assignment: `pitch ~ 1 + gender + context`. Let us look at it more closely

```{r, warning=FALSE, message=FALSE}
### Packages we will use ###
library(readr)
library(dplyr)
library(ggplot2)
### ### ###

df <- read_csv("https://tinyurl.com/polite-data") #download data
glimpse(df) #get a look at the data
```

We focus on the `pitch` (outcome / $y$) as well as the `gender` (predictor 1) and `context` (predictor 2) columns in this session. 

Let's again first visualize the data to see whether it even makes sense to predict pitch based on gender:

```{r}
ggplot(df, aes(x = pitch, fill = gender)) +
         geom_histogram(bins=70, alpha=0.6, position='identity') +
         theme_minimal() 

```


And the same for context:

```{r}
ggplot(df, aes(x = pitch, fill = context)) +
         geom_histogram(bins=50, alpha=0.6, position='identity') +
         theme_minimal() 
```
This is a more mixed picture than `gender`. We may also want to numerically see if there may be a difference in the mean pitch in formal vs. polite contexts:



```{r}
mean_informal <- df %>% 
                    filter(context == 'inf') %>%
                    pull() %>%
                    mean()
mean_polite <- df %>% 
                    filter(context == 'pol') %>%
                    pull() %>%
                    mean()

mean_informal
mean_polite
mean_informal - mean_polite
```

So there is a (numeric) difference but it is not clear whether it is meaningful.^[This is crucial to statistical analysis: What does a meaningful difference --in the context of your research question-- look like?] We will use a regression model to answer this question.^[Note that now we did not re-code `gender` and `context` using $1$ and $0$ as we did in the previous session. We will let R do this automatically for us.]

```{r}

pitch_model2 <- lm(formula = pitch ~ 1 + gender + context,
                   data    = df) 
summary(pitch_model2)
```

## Interactions

Is it reasonable to assume that `gender` and `context` do not interact? If we think that, maybe, the effect of `context` varies depending on the `gender` then we can add an interaction to the model. In formulas:
$$y = \beta_0 + \beta_\text{gender} \; x_{gender} + \beta_\text{context} \; x_{context} + \beta_\text{gender:context} \; x_{gender} \; x_{context}$$
Visually, 
```{r}
ggplot(data = df, aes(x = gender, y = pitch, fill = context)) +
  geom_boxplot() + 
  theme_minimal()
```
And the model itself,

```{r}

pitch_model3 <- lm(formula = pitch ~ 1 + gender * context,
                   data    = df) 
summary(pitch_model3)
```


# Model performance

## Goodness of fit
For example, $R^2$

## Predictive performance
For example, accuracy

## Corrections {.appendix}
If you spot a mistake or have suggestions, please get in touch with [Thomas Brochhagen](https://brochhagen.github.io) or [create an issue](https://github.com/metodesempirics/metodesempirics.github.io/issues/new)
