---
title: "Symbols and functions from today's session"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R 

### Naming (assigning a variable)

```{r, eval=FALSE}
any_name_I_like <- ... #gives "..." the name 'any_name_I_like'
```

For example,
```{r}
any_name_I_like <- 5 #now the value 5 is given the name 'any_name_I_like'
any_name_I_like + any_name_I_like #the same as 5+5
```

```{r}
another_name <- c(172,175,185) #"another_name"corresponds to a collection of three values
sum(another_name) #the sum of the values that go by the name "another_name"
```

### Mean

```{r, eval=FALSE}
mean(...) #computes the mean of "..."
```

For example,
```{r}
mean(c(172,175,185)) #the mean of the collection of values 172, 175, 185
mean(another_name)   #the mean of the values assigned to "another_name"
```

### Median

```{r, eval=FALSE}
median(...) #computes the median of "..."
```

For example,
```{r}
median(c(172,175,185)) #the median of the collection of values 172, 175, 185
median(another_name)   #the median of the values assigned to "another_name"
```

### Variance

```{r, eval=FALSE}
var(...) #computes the variance of "..."
```

For example, 
```{r}
var(c(172,175,185)) #the variance of the collection of values 172, 175, 185
var(another_name)   #the variance of the values assigned to "another_name"
```

### Standard deviation

```{r, eval=FALSE}
sd(...) #computes the standard deviation of "..."
```

For example,
```{r}
sd(c(172,175,185)) #the standard deviation of the collection of values 172, 175, 185
sd(another_name)   #the standard deviation of the values assigned to "another_name"
```

### Rounding

```{r, eval=FALSE}
round(..., digits = n) #round the value of "..." by n-digits after the comma
```

For example,
```{r}
round(123.4578, digits = 3)   #rounds 123.4578 to the third place after the comma
round(123.4578, digits = 1)   #rounds 123.4578 to the first place after the comma
```


### Mathematics
Capital sigma ($\sum$) is the *summation sign*. It's a convenient way to say that you will sum over all the values to the right of the sign. For instance,

$$\Large \sum_i x_i$$
means "sum over all the values in $x$"; $x_i$ refers to the observation $i$ in $x$. In other words, 

$$\Large \sum_i x_i = x_1 + ... + x_n$$,

where $x_1$ is the first element in $x$ and $x_n$ is the last. For example, if you have the values `172, 175` and `178` then your $x$ is a list of these values. 

$$\Large \sum_i x_i = 172 + 175 + 178$$

In R, this is the same as:

```{r}
x <- c(172, 175, 178) #I call my list of three observations x
sum(x)                #I sum the elements in x
```

which is the same as:
```{r}
172 + 175 + 178
```

Now what does, e.g., 
$$\Large \sum_i x_i^2$$ 
mean? 

It means that it is the sum of the elements of $x$ but squared. So $172^2 + 175^2 + 178^2$. In other words, apply everything that happens to the right of $\sum$ to each element of $x$ and then sum them up.

In R:
```{r}
x_squared <- x**2 #square all the elements in x and call them "x_squared"
sum(x_squared)    #the sum of the elements in "x_squared"

## which is the same as:
172**2 + 175**2 + 178**2
```

As a final example,

$$\Large \sum_i x_i + 1$$

takes each element in $x$, sums $1$ to it, and then adds them all up; so $(172 + 1) + (175 + 1) + (178 + 1)$

In R:
```{r}
x_plus_one <- x + 1 #add the value "1" to all elements in R
sum(x_plus_one)     #the sum of all elements in "x_plusöne"

## which is the same as:
(172 + 1) + (175 + 1) + (178 + 1)