---
title: "9 Visualization"
author:
- first_name: Thomas
  last_name: Brochhagen
  url: https://tbrochhagen.github.io/
  affiliation: Universitat Pompeu Fabra
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
  pdf_document:
    toc: yes
---


```{r preamble, echo=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(ggpubr)
```

# Visualization

Visualization is a key tool in empirical studies. The way we represent information can either help in conveying, or obfuscate, what we are trying to say. 

The field of visualization is still relatively new. Notwithstanding, there are some general guidelines that can help you along the way (see, e.g.,(Graphic Principles)[https://github.com/GraphicsPrinciples/CheatSheet/blob/master/NVSCheatSheet.pdf]:

* The "Why": Identify the purpose of your visual device (e.g., to make a contrast between two groups clear; or to visualize how the data is distributed)

* The "What": Identify the quantitative evidence to support the purpose

* The "Who": Identify the intended audience and design your visualization to aid them

* The "Where": Adapt design to presentation constraints


# Common types of visualizations

## Scatter plots

```{r, warning=FALSE, message=FALSE, echo=FALSE}

df <- read.csv("https://tinyurl.com/polite-data") #download data
```

```
## Corrections {.appendix}
If you spot a mistake or have suggestions, please get in touch with [Thomas Brochhagen](https://brochhagen.github.io) or [create an issue](https://github.com/metodesempirics/metodesempirics.github.io/issues/new)
