---
title: "4 Introduction to regressio analysis"
date: "`r Sys.Date()`"
author:
  - first_name: "Thomas"
    last_name: "Brochhagen"
    url: https://tbrochhagen.github.io/
    affiliation: Universitat Pompeu Fabra
bibliography: ../../biblio.bib
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
  pdf_document:
    toc: true
---



```{r preamble, echo=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(ggpubr)
```

## Regression
Regression is a fundamental technique for predicting an outcome variable from one or more predictors. For instance, the average pitch of a speaker (the outcome) based on their gender (predictor 1) and the context of interaction (predictor 2);  whether they will use a pronoun to refer to an entity (outcome) based on how predictable that entity is (predictor); or whether a linguistic school reform (predictor) increases the fluency of students in a language.

As summarized in  Gelman et al. [-@ros], some important uses of regression are:
  - prediction
  - exploring associations
  - extrapolation
  - causal inference

### Linear regression
The simplest regression model relates predictors and the outcome in a linear fashion: a linear regression model. With $x$ as a single predictor of $y$, we can write the model as:

$$y = \beta_0 + \beta_1x + \text{error} $$

In words, outcome $y$ is modelled as a linear combination of two parameters, $\beta_0$ and $\beta_1$, and an error term. The contribution of $\beta_1$ to the prediction is relative to the value of $x$, the predictor. 

Let us explore these ideas with a concrete example [@franke+roettger:20], with study predicting voice pitch (outcome / $y$) based on whether the speaker is male or female ($x$). We will add further predictors later.

#### Case study: pitch
```{r, warning=FALSE, message=FALSE}

```


## References

## Corrections {.appendix}
If you spot a mistake or have suggestions, please get in touch with [Thomas Brochhagen](https://brochhagen.github.io) or [create an issue](https://github.com/metodesempirics/metodesempirics.github.io/issues/new)
